---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "# ",
  fig.path = "figures/README-",
  dpi = 250
)
```


# __betalu__

__betalu__ implements the `(d/p/q/r)` statistics functions for the [general support beta distribution](https://en.wikipedia.org/wiki/Beta_distribution) (also called the four parameter beta) in [R](http://cran.r-project.org).  It is ideal for using in other packages since it is light weight and leverages the `(d/p/q/r)beta()` line of functions maintained by CRAN.



### Getting __betalu__

<!-- There are two ways to get __betalu__.  For the [CRAN version](https://cran.r-project.org/package=betalu), use -->
<!-- ```{r, eval=FALSE} -->
<!-- install.packages("betalu") -->
<!-- ``` -->
<!-- For the development version, use -->
Right now, you have to get __betalu__ from GitHub, since it's not on CRAN yet.  Here's how you do that:
```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("dkahle/betalu")
```



### The `(d/p/q/r)betalu()` functions

The defining property of the betalu distribution is that it is the beta distribution drawn to a different support.  In other words, it is the distribution of the random variable $Y = (u-l)*X + l$, where $X$ has a beta distribution.


The [PDF](https://en.wikipedia.org/wiki/Probability_density_function) (the _f(x)_) can be evaluated with the `dbetalu()` function:
```{r, fig.height = 3}
library(betalu)
library(ggplot2); theme_set(theme_bw())
x <- seq(-1, 1, .01)
qplot(x, dbetalu(x, 5, 5, -1, 1), geom = "line")
```
The distribution is the shifted and scaled version of the standard beta distribution with support [0,1], shown in red below:
```{r, fig.height = 3}
qplot(x, dbetalu(x, 5, 5, -1, 1), geom = "line") +
  stat_function(fun = dbeta, args = list(shape1 = 5, shape2 = 5), color = "red")
```


The [CDF](https://en.wikipedia.org/wiki/Cumulative_distribution_function) can be evaluated with the `pbetalu()` function:
```{r}
f <- function(x) dbetalu(x, 5, 5, -1, 1)
q <- -0.5
integrate(f, -1, q)
(p <- pbetalu(q, 5, 5, -1, 1))
```

The [quantile function](https://en.wikipedia.org/wiki/Quantile_function) can be evaluated with `qbetalu()`:
```{r}
qbetalu(p, 5, 5, -1, 1) # = q
```

And random number generation can be performed with `rbetalu()`:
```{r}
set.seed(1)
rbetalu(5, 5, 5, -1, 1)
```
```{r}
qplot(rbetalu(1e5, 5, 5, -1, 1), bins = 30)
```
`rbetalu()` can be used to obtain a [Monte Carlo](https://en.wikipedia.org/wiki/Monte_Carlo_method) estimate of the probability given by `pbetalu()` above:
```{r}
samples <- rbetalu(1e5, 5, 5, -1, 1)
mean(samples <= q)
```
Moreover, we can check the consistency and correctness of the implementation with
```{r, fig.height = 3}
qplot(samples, geom = "density") + 
  stat_function(fun = f,  color = "red")
```


